---
title: 'Reproducible Research: Peer Assessment 1'
output:
  html_document:
    keep_md: yes
  pdf_document: default
---


## Loading and preprocessing the data

1. Load the data

```{r echo=TRUE, message=FALSE, warning=FALSE}
library(dplyr)
unzip("activity.zip")
file <- read.csv("activity.csv")
file2 <- read.csv("activity.csv")
```

2. Process/transform the data (if necessary) into a format suitable for your analysis

```{r echo=TRUE, message=FALSE, warning=FALSE}
file$date <- as.Date(file2$date, "%Y-%m-%d")
file2$date <- as.Date(file$date, "%Y-%m-%d")
head(file)
```


## What is mean total number of steps taken per day?

1. Calculate the total number of steps taken per day

```{r echo=TRUE, message=FALSE, warning=FALSE}
steps_date <- file %>% group_by(date) %>% summarise("steps" = sum(steps, na.rm = TRUE))
print(steps_date)
```


2. Make a histogram of the total number of steps taken each day

```{r echo=TRUE}
hist(steps_date$steps, xlab = "Steps", main = "Frequency of Total Steps per Day", col = "rosybrown1")
```

3. Calculate and report the mean and median of the total number of steps taken per day

```{r echo=TRUE, message=FALSE, warning=FALSE}
report <- data.frame("mean" = mean(steps_date$steps, na.rm = TRUE),  "median" = median(steps_date$steps, na.rm = TRUE))
print(report)
```


## What is the average daily activity pattern?

1. Make a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r echo=TRUE, message=FALSE, warning=FALSE}
steps_interval <- file %>% group_by(interval) %>% summarise("steps" = mean(steps, na.rm = TRUE))
print(steps_interval)
```

```{r echo=TRUE}
plot(steps_interval$interval, steps_interval$steps, type = "l", ylab = "Steps", xlab = "Interval", main = "Avarage Steps per Interval", col = "rosybrown", lwd = 2)
```

2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r echo=TRUE, message=FALSE, warning=FALSE}
max_interval <- steps_interval[steps_interval$steps == max(steps_interval$steps),]
print(max_interval)
```


## Imputing missing values

1. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)

```{r echo=TRUE}
data_na <- sum(is.na(file))
print(data_na)
```

2. Devise a strategy for filling in all of the missing values in the dataset + 3. Create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r echo=TRUE, message=FALSE, warning=FALSE}
cases <- unique(file2$interval)
for (i in cases) { file2[(is.na(file) & file2$interval == i)] = steps_interval[(steps_interval$interval == i),2]}
head(file2)
```

4. Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day.

```{r echo=TRUE, message=FALSE, warning=FALSE}
steps_date2 <- file2 %>% group_by(date) %>% summarise("steps" = sum(as.numeric(steps)))
print(steps_date2)
```


```{r echo=TRUE}
hist(steps_date2$steps, xlab = "Steps", main = "Frequency of Total Steps per Day (fixed NAs)", col = "rosybrown1")
```

```{r echo=TRUE, message=FALSE, warning=TRUE}
report2 <- data.frame("NA" = c("corrected", "not corrected"),"mean" = c(mean(steps_date2$steps), mean(steps_date$steps, na.rm = TRUE)), "median" = c(median(steps_date2$steps), median(steps_date$steps, na.rm = TRUE)))
print(report2)
```


## Are there differences in activity patterns between weekdays and weekends?

1. Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.

```{r echo=TRUE, message=FALSE, warning=FALSE}
file2$day <- weekdays(file2$date, abbreviate = TRUE)
#abbreviations are in portuguese because of my configurations
file2$day[file2$day == "seg" | file2$day == "ter" | file2$day == "qua" | file2$day == "qui" | file2$day == "sex"] = "weekday"
file2$day[file2$day == "sáb" | file2$day == "dom"] = "weekend"
head(file2)
```


2. Make a panel plot containing a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).

```{r echo=TRUE, message=FALSE, warning=FALSE}
steps_day <- file2 %>% group_by(day, interval) %>% summarise("steps" = mean(as.numeric(steps)))
head(steps_day)
```

```{r echo=TRUE}
with(subset(steps_day, day == "weekday"), plot(interval, steps, type = "l", col = "brown1", main = "Avarage of steps per day", xlab = "Interval", ylab = "Steps"))
with(subset(steps_day, day == "weekend"), points(interval, steps, type = "l", col = "lightblue 4"))
legend("topright", legend = c("weekday", "weekdend"), col = c("brown1","lightblue3"), lwd= 2)
```

